---
title: "Malawi"
output: 
  html_document:
    toc: true
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE )

```

 
```{r packages, include= FALSE}

# install.packages("fabletools", repos = "https://tidyverts.org")
library( fabletools )
library( tsoutliers )
library( brolgar )
library( tsoutliers )
library( tsbox )
library( CausalImpact ) # loads MASS, which causes conflict with
library( anomalize )
library( tibbletime )
library( dygraphs )
library( tsbox )

library( readxl )

library( fpp3 ) # tsibbledata, feasts, fable, tsibble all-in one
library( GGally )
library( plotly )
library( sugrrants )

# Parallelization
# library(furrr)
# library( future.apply )
# plan(multiprocess)

library( tidyverse)
library( scales )
```

# Declining Confirmed Malaria Cases among children <5, Malawi, 2019
 
## data
 
 !Re-download formulas -- should have all elements from all formulas ...
 
```{r, datasets}

folder = '../dataDictionary/dhis2_dictionary/Formulas/'


formulas.file = 'Malawi_Formulas_2020-01-28.xlsx'
formulas = read_excel( paste0( folder, formulas.file ) , "Formula Elements")


confCases.file = enc2utf8('Malawi_confirmed cases _5yrs opd and chw_2020-01-28.xlsx')
stockout.file = 'Malawi_RDT Stock outs_2020-01-28.xlsx'
attendance.file = 'Malawi_OPD Attendance_2020-01-27.xlsx'

confCases.e = read_excel( paste0( folder, confCases.file ) , "formulaData")

confCases = read_excel( paste0( folder, confCases.file ) , "summaryData") %>%
  rename( orgUnit = orgUnit.x ) %>%
  mutate( confCases = ifelse( Count.max == 0 , 
                              NA, 
                              as.integer( sum )  )) %>%
  select( orgUnit, orgUnitName , period, confCases)

stockout = read_excel( paste0( folder, stockout.file ) , "formulaData")
attendance = read_excel( paste0( folder, attendance.file ) , "formulaData")

# Rainfall


# count( confCases.e , dataElement, Categories )

# count( stockout , dataElement, Categories )
# count( attendance , dataElement, Categories )
```

```{r, dataset}

# Drop count--because facility level data, and pivot 
idCols = c( 'orgUnit', 'orgUnitName' , 'period' )
nameCols = c('dataElement' , 'Categories' )
valueCols = 'SUM'

confCasesW = confCases.e %>% 
    mutate( SUM = as.integer( SUM ) ) %>%
    pivot_wider( all_of(idCols) , names_from = nameCols , values_from = valueCols )

stockoutW = stockout %>% 
    mutate( SUM = as.integer( SUM ) ) %>%
    pivot_wider( all_of(idCols) , names_from = nameCols , values_from = valueCols )

attendanceW = attendance %>% 
    mutate( SUM = as.integer( SUM ) ) %>%
    pivot_wider( all_of(idCols) , names_from = nameCols , values_from = valueCols )


# colnames( confCases )

dataset = confCasesW %>% 
    full_join( confCases , by = idCols ) %>%
    full_join( stockoutW , by = idCols ) %>%
    full_join( attendanceW , by = idCols ) 

# glimpse( dataset )
```
 - 41,000 observations of 10 variables identified by facility (orgUnit) and month (period)
 
```{r datasetTS}

datasetTS = dataset %>%
    mutate( month = zoo::as.yearmon( period, "%Y%m") %>%
                    yearmonth( . ) ) %>% 
    as_tsibble( ., key = c(orgUnit, orgUnitName) , index = month )

```

## Plotting TS 

```{r, spaghetti}

d = datasetTS %>% 
  # filter( orgUnit %in% atLeast48.confCases ) %>%
  select( orgUnit, month, confCases ) %>%
  as_tibble() %>%
  group_by( orgUnit ) %>%
  mutate_at( vars( confCases ) ,  
              list( confCases.scaled = ~  ifelse( row_number() == 1 | .x == 0 , 0, 
                                                  (.x - first(.x)) / first(.x) ) ,
                    confCases.logDiff = ~ log( .x ) - log( first(.x) )) 
             ) %>%
  as_tibble() 

g = d %>%
  ggplot() +
  geom_line( aes( x = month , y = confCases ,  
                  group = orgUnit) , alpha = .05
             )  

g %>% plotly::ggplotly()

g.scaled = d %>%
  ggplot() +
  geom_line( aes( x = month , y = confCases.logDiff ,  
                  group = orgUnit) , alpha = .05 
             )  

g.scaled %>% plotly::ggplotly()
```
 
 
## Annual bar chart

```{r national_annualized_bar_chart}

datasetTS.0 =  datasetTS %>% 
   summarise_if( is.integer , sum , na.rm = TRUE ) %>%
   mutate( orgUnit = 'O' , orgUnitName = 'National' ) 

national_annual = datasetTS.0 %>%
  as_tibble() %>%
  group_by( year = year(month) ) %>%
  summarise_if( is.integer , sum , na.rm = TRUE ) %>%
  as_tsibble( index = year )


```

# 2019 vs 2018

```{r annualized_2019_v_2018}

national_annual %>% 
  filter( year %in% 2018:2019 ) %>%
  ggplot( ) +
  geom_col( aes( x = year , y = confCases )) +
  scale_x_continuous( breaks = 2018:2019 , 
                      limits = c(2017,2020) ) +
  scale_y_continuous(labels = unit_format(unit = "M" , 
                                          scale = 1e-6)
                     ) +
  theme_bw() +
  theme( panel.grid.major = element_blank(),
            panel.grid.minor = element_blank() 
            ) +
  labs( title =  'Following Mass Bednet Distribution, Malaria Cases Declined in 2019' ,
        subtitle =  'Confirmed malaria cases in children <5 yrs of age' ,
        caption = 'Cases diagnosed by health facilities or community health workers' , 
        x = 'Year' , 
        y = 'Million Cases' ) 
```


```{r annualize_since_2015 }

national_annual %>% 
  ggplot( ) +
  geom_col( aes( x = year , y = confCases )) +
  scale_x_continuous( "" ,
                      breaks = 2015:2019 , 
                      limits = c(2014,2020) ) +
  scale_y_continuous( labels = unit_format(unit = "M" , 
                                          scale = 1e-6)
                     ) +
  theme_bw() +
  theme( panel.grid.major = element_blank(),
         panel.grid.minor = element_blank() ,
         text = element_text(size = 20)
            ) +
 labs( title =  'Following Mass Bednet Distribution, Malaria Cases Declined in 2019' ,
        subtitle =  'Confirmed malaria cases in children <5 yrs of age' ,
        caption = 'Cases diagnosed by health facilities or community health workers' , 
        # x = 'Year' , 
        y = 'Cases (Millions) ' ) + 
  geom_abline( aes( intercept = 
                      mean(national_annual$confCases) , slope = 0 ),
                   color = 'red' ) +
  annotate("text", label = "Mean", x = 2014.5, y = 2.8e6, 
           size = 10, colour = "red" )



```

## Monthly cases 

```{r national_monthly_confirmed_cases}

  datasetTS.0 %>% 
       autoplot( confCases   ) +
       labs( title = 'confCases' ) +
       guides( color = FALSE ) +
       theme_bw( base_size = 10 )


```


## Monthly cases with counts 

```{r national_monthly_confirmed_cases_counts}

cols = setdiff( names( datasetTS.0 ) , c( 'month' , 'orgUnit' , 'orgUnitName', 'var' ))


datasetTS.0.sum_count.monthly = 
  inner_join( 
  datasetTS.0.n %>% rename_at( vars(cols), ~ paste0(., "_count") ) , 
  datasetTS.0 , 
  by = c('month' , 'orgUnit', 'orgUnitName' )
  ) %>% 
  # mutate_if( is.numeric , scale ) %>%
  mutate(  year = year( month )  ) %>% 
  
  as_tibble() %>% 
  select(  month , year , orgUnit , starts_with( 'confCases' ) )   %>% 
  group_by( month , orgUnit = 'National' ) %>%
  summarise_if( is.numeric , sum ) %>%
  
  as_tsibble( index = month , key = orgUnit ) 

datasetTS.0.sum_count.monthly = 
  inner_join( 
  datasetTS.0.n %>% rename_at( vars(cols), ~ paste0(., "_count") ) , 
  datasetTS.0 , 
  by = c('month' , 'orgUnit', 'orgUnitName' )
  ) %>% 
  # select( month , orgUnit , starts_with( 'confCases' ) )   %>% 
  as_tibble() %>%
  group_by( month  ) %>%
  summarise_if( is.numeric , sum ) %>% 
  ungroup %>% 
  mutate_at( vars(-month ) , 
            ~ log( .x ) - log( first(.x) ) 
             ) %>%
  mutate( orgUnit = 'National' ) %>%
  
  as_tsibble( index = month  , key = orgUnit ) 

datasetTS.0.sum_count.monthly %>% 
       autoplot() +
       labs( title = .x ) +
       guides( color = FALSE ) +
       theme_bw( base_size = 10 )


datasetTS.0.sum_count.monthly  %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))


```

```{r annualized_scaled_confCases }


datasetTS.0.sum_count.annual = 
  inner_join( 
  datasetTS.0.n %>% rename_at( vars(cols), ~ paste0(., "_count") ) , 
  datasetTS.0 , 
  by = c('month' , 'orgUnit', 'orgUnitName' )
  ) %>% 
  # mutate_if( is.numeric , scale ) %>%
  mutate(  year = year( month )  ) %>% 
  
  as_tibble() %>% 
  select( year , orgUnit , starts_with( 'confCases' ) )   %>% 
  group_by( year , orgUnit = 'National' ) %>%
  summarise_if( is.numeric , sum ) %>% 
  ungroup %>% 
    ungroup %>%
  mutate_at( vars(-year, -orgUnit) , 
            ~ log( .x ) - log( first(.x) )
             ) %>%
  
  as_tsibble( index = year  , key = orgUnit ) 


datasetTS.0.sum_count.annual  %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))
```


```{r datasetTS.0.sum_count_adjusted }
maxn = max( datasetTS.0.sum_count$confCases_count )

datasetTS.0.sum_count %>% 
  ungroup %>% 
  mutate( 
    adj.n = maxn / confCases_count  , 
    adj =  confCases * adj.n  ) %>%
  select( orgUnit, year, confCases , adj ) %>%
  # mutate_at( vars(-year, -orgUnit) , scale ) %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))
```


```{r annualized_scaled }

cols = setdiff( names( datasetTS.0 ) , c( 'month' , 'orgUnit' , 'orgUnitName'))

# scaled
national_annual %>%
  mutate_at( cols , ~ log( .x ) - log( first(.x) ) ) %>%
  # select(  year, confCases ) %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))
```


## Seaonality of data

Season plot (quarterly), aggregated data

```{r, national_trends_facet_seasonal }

endpoints = datasetTS.0 %>% 
    mutate( year = lubridate::year( month ) , 
            mnth = lubridate::month( month ) ) %>%
    filter( mnth == 12 )

datasetTS.0 %>% gg_season( confCases ) +
         ggtitle( 'confCases' ) + 
         theme_bw( base_size = 10 )


datasetTS.0 %>% as_tibble() %>%
  group_by( q = yearquarter( month) ) %>%
  summarise_if( is.integer, sum , na.rm = TRUE )  %>%
  as_tsibble( index =  q ) %>%
  fill_gaps() %>%
  # select( -month ) %>%
  gg_subseries( confCases ) +
         ggtitle( 'confCases' ) + 
         theme_bw( base_size = 10 )

```
 
 



# National model 


There were fewer confirmed malaria cases in 2019. Restricted analyses to confirmed malaria cases in children < 5. 


```{r, national_count_trends_facet }

datasetTS.0.n =  datasetTS %>% 
   summarise_if( is.integer , ~sum(!is.na(.x))  ) %>%
   mutate( orgUnit = 'O' , orgUnitName = 'National' ) 

cols = setdiff( names( datasetTS.0 ) , c('orgUnit', 'orgUnitName' , 'month'))

# map( cols , ~datasetTS.0 %>% autoplot( vars( !!.x )   ) +
#          ggtitle( .x ) + theme_bw( base_size = 10 )
#      )


map( cols , ~bind_rows( 
  datasetTS.0.n %>% mutate( var = 'Count' ) %>% as_tibble() ,
  datasetTS.0 %>% mutate( var = 'Sum' ) %>% as_tibble()
                        )  %>% 
       as_tsibble( index = month, key = c( var, orgUnit, orgUnitName ) ) %>%
       autoplot( vars( !!.x )   ) +
       facet_grid( var  ~ . , scale = 'free_y' ) +
       expand_limits( y = 0 ) +
       labs( title = .x ) +
       guides( color = FALSE ) +
       theme_bw( base_size = 10 )
     )

```

## Frequency of monthly data submitted by facility 

how many facilities have enough data for modeling time-series? 

```{r, histogram_monthly_submissions}

orgUnit.counts = 
  dataset %>%
  group_by( orgUnit ) %>% 
  summarise_if( is.integer, ~sum( !is.na(.) ) ) 

confCases.counts = orgUnit.counts %>%
  select( orgUnit , confCases ) %>%
  mutate( years_of_data = ( confCases / 12 ) %>% round() )  

map( cols , 
     ~ orgUnit.counts %>%
       # select( orgUnit , .x ) %>%
       mutate( years_of_data = ( !! rlang::sym( .x ) / 12 ) %>% 
                 round(0) %>% as.integer())  %>%
       ggplot( aes( years_of_data ) ) +
       geom_histogram( binwidth = 1 ) +
       labs( title = 'Frequency of monthly data available for each orgUnit' ,
             subtitle = paste( .x , "(" , 
                               nrow( orgUnit.counts ) , "facilities )" ) ,
             y = 'Number of Facilities' ,
             x = 'Years of Data' 
  ) +
    expand_limits( x = 5 ) +   
    stat_bin( binwidth=1 , geom="text", colour="white", size=3.5,
           aes( label = ..count.. ), 
           position=position_stack( vjust=0.5) ) +
    theme_bw()
)

# Facility has at least 4 years of confCases daata
atLeast4Years = orgUnit.counts %>% 
  group_by( orgUnit ) %>% 
  summarise_at( cols , ~ ( .x / 12 ) %>% round() >= 4 ) 

atLeast4Years.confCases = atLeast4Years %>% 
  filter( confCases == TRUE ) %>% 
  pull( orgUnit )

# Facility has no missing data

maxMonths = length( unique( dataset$period ) )  

noMissing = orgUnit.counts %>% 
  group_by( orgUnit ) %>% 
  summarise_at( cols , ~ .x == maxMonths ) 

noMissing.confCases = noMissing %>% 
  filter( confCases == TRUE ) %>% 
  pull( orgUnit )

# Bins of confCases data
confCases.n = 
  dataset %>%
  group_by( orgUnit ) %>% 
  summarise_at( vars(confCases), ~sum( !is.na(.) ) ) %>%
  mutate( n = cut( confCases , breaks = 5) )

confCases.n %>% ggplot() + geom_histogram( aes(n) , stat = 'count')

atLeast48.confCases = confCases.n %>%
  filter( confCases >= 48 ) %>%
  pull( orgUnit )

length( atLeast48.confCases )

```


## Summarise missingness

```{r missingness }

```

### Number of facilities with complete data for each series.

### Could reporting be biased by location and/or RDT stockouts? Seasonal plot of reporting.  


Combined series, showing number complete.  



## Seasonality of missing data? 

```{r}

datasetTS.0.n %>% gg_season( confCases ) +
  labs( title = 'Count of facilities submitting confCases data') +
  theme_bw( base_size = 10 )


datasetTS.0.n %>% gg_subseries( confCases ) +
  labs( title = 'Count of facilities submitting confCases data') +
  theme_bw( base_size = 10 )

datasetTS.0.n %>% as_tibble() %>%
  group_by( q = yearquarter( month) ) %>%
  summarise_if( is.integer, mean , na.rm = TRUE )  %>%
  as_tsibble( index =  q ) %>%
  fill_gaps() %>%
  # select( -month ) %>%
  gg_subseries( confCases ) +
         ggtitle( 'Mean count of facilities submitting confCases data each month, by quarter' ) + 
         theme_bw( base_size = 10 )

maxSubmissions.confCases = datasetTS.0.n %>% pull( confCases ) %>% max 
maxSubmissionsMonth.confCases = datasetTS.0.n %>%  
  filter( confCases == maxSubmissions.confCases ) %>%
  pull( month )

cat( 'The greatest number of confCases submissions in any one month was' ,
     comma(maxSubmissions.confCases),  'in' , 
     paste( maxSubmissionsMonth.confCases , collapse = ', ') 
     )

```


## Impact, aggregated data, national level.  Naive, or crude, model assumes things that probably should not be assumed without good evidence.  

```{r aggregated_features}

# Transform??? 

datasetTS.0 %>% autoplot( confCases )
datasetTS.0 %>% autoplot( difference( confCases , 2) )

datasetTS.0 %>% autoplot( log( confCases) )
datasetTS.0 %>% autoplot( difference( log(confCases) , 2) )

lambdaG <- datasetTS.0 %>%
  features( confCases , features = guerrero) %>%
  pull(lambda_guerrero)

datasetTS.0 %>% autoplot( box_cox( confCases, lambdaG ) )
datasetTS.0 %>% autoplot( difference( box_cox( confCases, lambdaG ) , 2) )

# no transformation
datasetTS.0 %>% 
  mutate( 
    x = confCases ,
    diff12 = difference( lag(x, 1) , 12 )  ) %>% 
  gg_tsdisplay( diff12 , plot_type='partial') +
  ggtitle( 'no transformation' )

# log 
datasetTS.0 %>% 
  mutate( 
    x = log( confCases + 1 )  ,
    diff12 = difference( lag( x , 1) , 12 )  ) %>% 
  gg_tsdisplay( diff12 , plot_type='partial') +
  ggtitle( 'log transformation' )

# Lamba guerrero()
datasetTS.0 %>% 
  mutate( 
    x = box_cox( confCases, lambda = lambdaG ) ,
    diff12 = difference( lag( x , 1) , 12 )  
    ) %>% 
  gg_tsdisplay( diff12 , plot_type='partial') +
  ggtitle( 'lambda guerrero transformation' )

# Seasonal? 
datasetTS.0 %>% autoplot( difference( confCases , 2) )

datasetTS.0 %>% gg_subseries( confCases)

datasetTS.0 %>% gg_lag 

datasetTS.0 %>% gg_tsdisplay( confCases , plot_type='partial')

datasetTS.0 %>% 
  mutate( 
    x = confCases ,
    diff12 = difference( lag(x, 1) , 12 )  ) %>% 
  gg_tsdisplay( diff12 , plot_type='partial') +
  ggtitle( 'no transformation' )

```

## cross-validation (national)

```{r crossValidation0}

# https://otexts.com/fpp3/tscv.html
# sction 5.7 

train.0.cross = datasetTS.0 %>%
    filter(
        month < yearmonth( "2018-Dec" )
    ) %>%
    select( orgUnit, month, confCases ) %>%
    slide_tsibble( .size = 36 , .step = 1 )


fit.train.0.cross = train.0.cross %>% model( model_list )

    model( 

          ets_log = ETS( log(confCases + 1)  ) ,
          
          etsMNA_log = ETS( log(confCases + 1) ~
                          error('M') + trend('N') + season('A') ) ,
          
          etsANA_log = ETS( log(confCases + 1) ~
                          error('A') + trend('N') + season('A') ) ,

          arima_log = ARIMA( log(confCases + 1) ~ 
                           pdq(0,1,1) + PDQ(0,1,0)) ,
          
          arima_log_auto = ARIMA( log(confCases + 1) ~ 
                           pdq() + PDQ()) ,

          fourier1_log = ARIMA( log(confCases + 1) ~
                              fourier( K = 1 ) + PDQ(0,0,0) ) , 
          
          ets = ETS( confCases + 1  )   ,
          
          etsMNA = ETS( confCases + 1   ~
                          error('M') + trend('N') + season('A') ) ,
          
          etsANA = ETS( confCases + 1  ~
                          error('A') + trend('N') + season('A') ) ,

          arima = ARIMA( confCases + 1  ~ 
                           pdq(0,1,1) + PDQ(0,1,0)) ,
          
          arima_auto = ARIMA(confCases + 1   ~ 
                           pdq() + PDQ()) ,

          fourier1 = ARIMA( confCases + 1 ~
                              fourier( K = 1 ) + PDQ(0,0,0))
           ) 

# check fit parameters
map(1:12 , ~fit.train.0.cross %>% 
      filter( .id == .x ) %>% select(arima_auto) %>% map(1))

map(1:12 , ~fit.train.0.cross %>% 
      filter( .id == .x ) %>% select(ets) %>% map(1))

map(1:12 , ~fit.train.0.cross %>% 
      filter( .id == .x ) %>% select(ets) %>% 
      components() %>%
  autoplot() +
  ggtitle( 'ets')
  )

# fore.train.0.cross  = fit.train.0.cross %>%  
#   select( -ets_damped_ad , -ets_damped_md) %>%
#   forecast( h = 1 )
# 
# fore.train.0.cross %>% accuracy( datasetTS.0 )

fc = fit.train.0.cross %>%  
  forecast( h=12 ) %>%
  group_by(.id , .model ) %>%
  mutate( h = row_number() ) %>%
  ungroup()

g = fc %>%
  accuracy( datasetTS.0 , by = c("h",".model") ) %>%
  ggplot(aes(x = h, y = RMSE, color = .model, 
             group = .model)) + 
  geom_line() 

plotly::ggplotly( g )

save( fc , train.0.cross, 
      fit.train.0.cross ,
      file = 'train0.cross.Rda' )

```

 
We selected 5 years of data for several variables of confirmed cases and potentially explanatory variables, including attendance.  (describe, zeros?)  
 

```{r aggregated_model} 

 # whole data set
fit0 = datasetTS.0 %>% 
    fill_gaps() %>%
    select( orgUnit, orgUnitName, month, confCases ) %>%

    model( ets = ETS(  log( confCases ) ) ,
           ets.damped = ETS( confCases ~ trend("Ad") ) ,
           stl = STL( confCases ~ season(window=12) + trend(window=11)) ,
           arima = ARIMA( log( confCases ) ) ,
           fourier = ARIMA( log( confCases ) ~ fourier(K = 1) + PDQ(0,0,0))
                                     )
# plot fiited values 

fit0 %>% 
  select( - stl ) %>%
  augment() %>% 
  autoplot(.fitted) +
  autolayer( datasetTS.0, confCases, color = 'black', alpha = .5  )

# fit0 %>% accuracy( )


```

Models for ETS and ARIMA run only when pre-conditioning data with log.  Is there a better conditioning factor? 

```{r aggregated_model_leave_out}

# trainin set
train.0 = datasetTS.0 %>% 
  filter( month < yearmonth( "2018-Dec" )  ) %>%
    select( orgUnit, month, confCases ) 

fit.train.0 = train.0 %>% 
    model( 

          ets = ETS( log(confCases + 1)  ) ,

          ets_damped = ETS( log(confCases + 1)  ~ trend("Md") ,
                            restrict = FALSE ) ,

          arima = ARIMA( log(confCases + 1) ~ 
                           pdq(0,1,1) + PDQ(0,1,0)) ,
          
          arima_auto = ARIMA( log(confCases + 1) ~ 
                           pdq() + PDQ()) ,


          fourier1 = ARIMA( log(confCases + 1) ~
                              fourier( K = 1 ) + PDQ(0,0,0)) 
           ) 
                                 

# fit.train.0 %>% components() %>% autoplot( season_adjust )

fit.train.0 %>% accuracy()

# fit.train.0 %>% features( confCases )

# report( fit.train.0 )

# forecast 
fore = fit.train.0 %>% 
    # select( orgUnit, ets )  %>% 
    forecast( h = '13 months' )

fore %>% 
  autoplot( level = NULL ) + 
  autolayer( datasetTS.0 , confCases ) 

accuracy( fore, datasetTS.0 )

map( c('ets', 'ets_damped', 'arima' , 'arima_auto' , 'fourier1'  ) , 
     ~fore %>% filter( .model %in% .x ) %>%  autoplot() + 
         autolayer( datasetTS.0 , confCases ) +
         ggtitle( .x )
)


```

For 2 of three models, the actual is within the 80% credible prediction interval.  We need to pay attention to which model is the best fit.  


## Crude reporting adjustment

```{r confCases_crude_adjust}

 
 max.confCases.n = datasetTS.0.n %>% pull( confCases ) %>% max(. , na.rm = TRUE )

datasetTS.0.adj = datasetTS.0 %>% 
  select( month , confCases ) %>%
  inner_join( datasetTS.0.n %>% select( month, confCases ) %>% 
                rename( count = confCases) , by = "month") %>%
  mutate( crude.adj = max.confCases.n / count  ,
          confCases.adj = confCases * crude.adj ) 

datasetTS.0.adj  %>% 
       gather("var" , 'val', -month, -count  ) %>%
       filter( !var %in% 'crude.adj' ) %>%
       autoplot( val  ) +
       # facet_grid( var  ~ . ,scale = 'free_y' ) +
       # labs( title = .x ) +
       # guides( color = FALSE ) +
       theme_bw( base_size = 10 )


```


# District model 


```{r ous, eval=FALSE}

meta = readRDS( '../HMIS/Surv_Assessment/Malawi/datasets/Malawi_metadata.rds' )

source( '../HMIS/Surv_Assessment/dhis2_functions.R')

ous = ous_from_metatdata( .meta = meta , 
                          simplify = TRUE , simplify.keep = .02 , SF = TRUE )

saveRDS( ous , file = 'malawi_ous.rda')

```


```{r districts }
ous = readRDS( 'malawi_ous.rds' )

districts = ous %>% filter( level ==3 ) %>%
  mutate( district = orgUnit.name ,
          district.id = orgUnit )

region_district = districts %>%
  select( orgUnit.name, parent_ou.name, district) %>%
  arrange( parent_ou.name, district ) 

# View( region_district )

datasetTS.3 = datasetTS %>% 
  left_join( ous %>% select( orgUnit, parent_ou ) ,
             by = 'orgUnit' ) %>%
  left_join( districts %>% 
               select( district, district.id ), 
             by = c( 'parent_ou' = 'district.id' )
  )  %>%
  filter( !is.na( district ) ) %>%
  group_by( district ) %>%
  summarise_if( is.integer , sum , na.rm = TRUE ) 


```


```{r district_absolute}

datasetTS.3 %>% autoplot( confCases  ) %>%
  plotly::ggplotly()


```

Lilongwe district is greatest contributor to number of confirmed cases 

```{r district_percent_total}

# examine contribution to the total each month 

datasetTS.3 %>% 
  pivot_longer( !matches( c('month', 'district') ) , names_to = 'var' , values_to = 'val' ) %>% 
  group_by( month , var ) %>% 
  mutate( total = sum( val, na.rm = TRUE )) %>%
  group_by( month, var ,district ) %>%
  mutate( pct = val / total ) %>%
  select( - total , -val ) %>%
  pivot_wider( names_from = 'var' , values_from = 'pct') %>% 
  as_tsibble( index = month , key = district ) %>%
  
  # Facility RDT
  select( month, district , contains('RDT') & contains('NMCP') ) %>%
  
  ts_xts( ) %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))
  


```

Lilongwe district acounts for 5-20% of cases each month.  3 districts acount for 30% each month (Lilongwe , Mangochi, Kasungu )



```{r district_scaled }

datasetTS.3 %>% 
  group_by( district  ) %>%
  mutate( confCases.scaled = scale( confCases ) ,
          confCases.logDiff = log( confCases) - log( first(confCases) ) ) %>%
  # autoplot( confCases.scaled   ) %>%
  select( district , month, confCases.logDiff ) %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))

```

When scaled,Mwanza Feb 2017 is 5 SD, whereas other series have no values >3sd 

When log difference, Chiradzulu has greatest change, ~ 3x


## Annualized cases

```{r annualized_districts}
district_annual = datasetTS.3 %>%
  as_tibble() %>%
  group_by( district , year = year(month) ) %>%
  summarise_if( is.integer , sum , na.rm = TRUE ) %>%
  as_tsibble( key = district , index = year )

district_annual %>% autoplot( confCases ) %>%
  ggplotly() %>% hide_legend()

# scaled
district_annual %>%
  mutate_at( cols , scale) %>%
  select( district , year, confCases ) %>%
  ts_xts() %>%
  dygraphs::dygraph( . ) %>%
  dyEvent("2018-11-15", "Net Campaign", 
          labelLoc = "top") %>%
  dyHighlight(
    highlightSeriesOpts = list(strokeWidth = 5)
    ) %>% 
  dyLegend(show = "follow") %>%
  dyCSS(textConnection("
     .dygraph-legend > span { display: none; }
     .dygraph-legend > span.highlight { display: inline; }
  "))

```

```{r, district_count_trends_facet }

datasetTS.3.n = datasetTS %>% 
  left_join( ous %>% select( orgUnit, parent_ou ) ,
             by = 'orgUnit' ) %>%
  left_join( districts %>% 
               select( district, district.id ), 
             by = c( 'parent_ou' = 'district.id' )
  )  %>%
  filter( !is.na( district ) ) %>%
  group_by( district ) %>%
  summarise_if( is.integer , ~sum(!is.na(.x))  ) 

cols = setdiff( names( datasetTS.3 ) , c( 'month' , 'district'))

# map( cols , ~datasetTS.0 %>% autoplot( vars( !!.x )   ) +
#          ggtitle( .x ) + theme_bw( base_size = 10 )
#      )


map( cols , ~bind_rows( 
  datasetTS.3.n %>% mutate( var = 'Count' ) %>% as_tibble() ,
  datasetTS.3 %>% mutate( var = 'Sum' ) %>% as_tibble()
                        )  %>% 
       as_tsibble( index = month, 
                   key = c( var, district ) ) %>%
       autoplot( vars( !!.x )   ) +
       facet_grid( var  ~ . ,scale = 'free_y' ) +
       labs( title = .x ) +
       guides( color = FALSE ) +
       theme_bw( base_size = 10 )
     )

```

## Frequency of monthly data submitted by facility 

how many facilities have enough data for modeling time-series? 

```{r, district_histogram_monthly_submissions}

orgUnit.counts = 
  dataset %>%
  group_by( orgUnit ) %>% 
  summarise_if( is.integer, ~sum( !is.na(.) ) ) 

confCases.counts = orgUnit.counts %>%
  select( orgUnit , confCases ) %>%
  mutate( years_of_data = ( confCases / 12 ) %>% round() )  

map( cols , 
     ~ orgUnit.counts %>%
       # select( orgUnit , .x ) %>%
       mutate( years_of_data = ( !! rlang::sym( .x ) / 12 ) %>% 
                 round(0) %>% as.integer())  %>%
       ggplot( aes( years_of_data ) ) +
       geom_histogram( binwidth = 1 ) +
       labs( title = 'Frequency of monthly data available for each orgUnit' ,
             subtitle = paste( .x , "(" , 
                               nrow( orgUnit.counts ) , "facilities )" ) ,
             y = 'Number of Facilities' ,
             x = 'Years of Data' 
  ) +
    expand_limits( x = 5 ) +   
    stat_bin( binwidth=1 , geom="text", colour="white", size=3.5,
           aes( label = ..count.. ), 
           position=position_stack( vjust=0.5) ) +
    theme_bw()
)

# Facility has at least 4 years of confCases daata
atLeast4Years = orgUnit.counts %>% 
  group_by( orgUnit ) %>% 
  summarise_at( cols , ~ ( .x / 12 ) %>% round() >= 4 ) 

atLeast4Years.confCases = atLeast4Years %>% 
  filter( confCases == TRUE ) %>% 
  pull( orgUnit )

# Facility has no missing data

maxMonths = length( unique( dataset$period ) )  

noMissing = orgUnit.counts %>% 
  group_by( orgUnit ) %>% 
  summarise_at( cols , ~ .x == maxMonths ) 

noMissing.confCases = noMissing %>% 
  filter( confCases == TRUE ) %>% 
  pull( orgUnit )

# Bins of confCases data
confCases.n = 
  dataset %>%
  group_by( orgUnit ) %>% 
  summarise_at( vars(confCases), ~sum( !is.na(.) ) ) %>%
  mutate( n = cut( confCases , breaks = 5) )

confCases.n %>% ggplot() + geom_histogram( aes(n) , stat = 'count')

atLeast48.confCases = confCases.n %>%
  filter( confCases >= 48 ) %>%
  pull( orgUnit )

length( atLeast48.confCases )

```

## Summarise missingness

```{r district_missingness }

```

### Number of facilities with complete data for each series.

### Could reporting be biased by location and/or RDT stockouts? Seasonal plot of reporting.  


Combined series, showing number complete.  


## Seaonality of data

Season plot (quarterly), aggregated data

```{r, district_trends_facet_seasonal }

endpoints = datasetTS.3 %>% 
    mutate( year = lubridate::year( month ) , 
            mnth = lubridate::month( month ) ) %>%
    filter( mnth == 12 )

datasetTS.3 %>% gg_season( confCases ) +
         ggtitle( 'confCases' ) + 
         theme_bw( base_size = 10 )


datasetTS.3 %>% as_tibble() %>%
  group_by( q = yearquarter( month) ) %>%
  summarise_if( is.integer, sum , na.rm = TRUE )  %>%
  as_tsibble( index =  q ) %>%
  fill_gaps() %>%
  # select( -month ) %>%
  gg_subseries( confCases ) +
         ggtitle( 'confCases' ) + 
         theme_bw( base_size = 10 )

```
 
## Seasonality of missing data? 

```{r district_seasonal_missing}

datasetTS.3.n %>% gg_season( confCases ) +
  labs( title = 'Count of facilities submitting confCases data') +
  theme_bw( base_size = 10 )


datasetTS.3.n %>% gg_subseries( confCases ) +
  labs( title = 'Count of facilities submitting confCases data') +
  theme_bw( base_size = 10 )

datasetTS.3.n %>% as_tibble() %>%
  group_by( q = yearquarter( month) ) %>%
  summarise_if( is.integer, mean , na.rm = TRUE )  %>%
  as_tsibble( index =  q ) %>%
  fill_gaps() %>%
  # select( -month ) %>%
  gg_subseries( confCases ) +
         ggtitle( 'Mean count of facilities submitting confCases data each month, by quarter' ) + 
         theme_bw( base_size = 10 )

maxSubmissions.confCases = datasetTS.3.n %>% pull( confCases ) %>% max 
maxSubmissionsMonth.confCases = datasetTS.3.n %>%  
  filter( confCases == maxSubmissions.confCases ) %>%
  pull( month )

cat( 'The greatest number of confCases submissions in any one month was' ,
     comma(maxSubmissions.confCases),  'in' , 
     paste( maxSubmissionsMonth.confCases , collapse = ', ') 
     )

```


## Model
Impact, aggregated data, national level.  Naive, or crude, model assumes things that probably should not be assumed without good evidence.  

```{r district_aggregated_features, eval = FALSE}

# Transform??? no
lambda <- datasetTS.3 %>%
  features( confCases , features = guerrero) %>%
  pull(lambda_guerrero)

datasetTS.3 %>% gg_subseries( confCases)

datasetTS.3 %>% gg_lag 

datasetTS.3 %>% autoplot( log( confCases) )
datasetTS.3 %>% autoplot( difference( log(confCases) , 2) )

datasetTS.3 %>% 
  mutate( diffl2 = difference( lag(confCases, 1) , 12 )  ) %>% 
  gg_tsdisplay( diffl2 , plot_type='partial')

datasetTS.3 %>% 
  mutate( diffl2 = difference( log(confCases) , 12 ) ) %>%
  ACF( diffl2) %>% autoplot()

  # features( diffl2, ljung_box, lag = 10)


```

```{r district_aggregated_model} 

 # whole data set
fit3 = datasetTS.3 %>% 
    fill_gaps() %>%
    select( district, month, confCases ) %>%
    model( 
      ets = ETS(  log( confCases + 1) ) ,
      ets.damped = ETS( confCases ~ trend("Ad") ) ,
      arima = ARIMA( log( confCases +1 ) ) ,
      fourier = ARIMA( log( confCases + 1) ~ fourier(K = 1) + PDQ(0,0,0))
                                 )

# plot fiited values 

fit3 %>% 
  # select( - stl ) %>%
  augment() %>% 
  autoplot(.fitted) +
  autolayer( datasetTS.3, confCases, color = 'black', alpha = .5  )

# fit3 %>% accuracy( )


```

Models for ETS and ARIMA run only when pre-conditioning data with log.  Is there a better conditioning factor? 

```{r district_aggregated_model_leave_out}

# trainin set
train.3 = datasetTS.3 %>% 
  filter( month < yearmonth( "2018-Dec" )  ) %>%
    select( district, month, confCases ) %>% 
    fill_gaps()

fit.train.3 = train.3  %>%
    model( 
      ets = ETS(  log( confCases + 1) ) ,
      ets.damped = ETS( confCases ~ trend("Ad") ) ,
      arima = ARIMA( log( confCases +1 ) ) ,
      fourier = ARIMA( log( confCases + 1) ~ fourier(K = 1) + PDQ(0,0,0))
                                 )

# fit.train.0 %>% components() %>% autoplot( season_adjust )

fit.train.3 %>% accuracy()

# fit.train.3 %>% features( confCases )

# report( fit.train.3 )

# forecast 
fore3 = fit.train.3 %>% 
    # select( orgUnit, ets )  %>% 
    forecast( h = '13 months' )

# fore3 %>% autoplot( level = NULL ) + 
#   autolayer( datasetTS.3 , confCases ) +
#   guides( color = FALSE )

accuracy( fore3, datasetTS.3 )

map( c('ets', 'ets.damped' , 'arima' , 'fourier') , 
     ~fore3 %>% filter( .model %in% .x ) %>%  
       autoplot() + 
       autolayer( datasetTS.3 , confCases ) +
       ggtitle( .x )
)


```



## cross-validation (district)

```{r crossValidation3_model}

train.3.cross = datasetTS.3 %>%
    select( district, month, confCases ) %>%
    stretch_tsibble(.init = 3, .step = 1)


fit.train.3 = train.3.cross %>% 
    model( 
      # rw = RW( confCases  ~ drift()) ,
           ets = ETS( log(confCases + 1 )  ) ,
           # etsadj = ETS( log(confCases + 1 )  ~ trend("Ad") ) ,
           # fourier1 = ARIMA( log(confCases + 1) ~ fourier(K = 1) + PDQ(0,0,0)) ,
           # arima = ARIMA( log( log(confCases + 1) ) )
           ) 

```

```{r crossValidation3_accuracy}
# fit.train.3 %>% accuracy()

# fit.train.3  %>% 
#   select( orgUnit , etsadj ) %>% 
#   components( fit.train.3 ) %>% autoplot()

fit.train.3  %>% glance() %>% group_by(.model) %>% 
  summarise_if( is.numeric, mean , na.rm = TRUE )

# fit.train.3 %>% select( .id, district, fourier1 ) 
# fit.train.3 %>% select( .id, district, arima ) 

fore.train.3  = fit.train.3 %>%  forecast( h = '1 year' )

fore.train.3 %>% accuracy( datasetTS.3 )

# Residual accuracy ? 
# datasetTS.3 %>% model(RW( confCases ~ drift())) %>% accuracy()

```

 
# Facility model  


```{r train4}
# trainin set
train.4 = datasetTS %>% 
  filter( 
    month < yearmonth( "2018-Dec" ) ,
    orgUnit %in% atLeast48.confCases
    ) %>% 
  select( orgUnit, month, confCases ) %>% 
  fill_gaps()

ou = train.4$orgUnit %>% unique()

```


## time series features 

```{r time-series-features, eval=FALSE}

i = which( ou =='eZfOjVQcd0e' )
## How to decipher wihich have a model???
i = 100:200
i = 1:length( ou )

# select by ou id
# i = which( ou %in% 'asiPyqS8rHs')
if ( i < 6 ){
    train.4 %>% filter( orgUnit == ou[i] ) %>% autoplot(confCases) +
  # labs( title = i ) + 
  guides( color = FALSE )
}

# time-series features
f = map( i , ~ train.4 %>% 
                filter( orgUnit == ou[.x] ) %>%
                features( confCases  , 
                          features = list( 
                              n = ~ sum( !is.na(.) ), 
                              mean = ~ mean(., na.rm = TRUE) ,
                              sd = ~ var(., na.rm = TRUE)^.5, 
                              # ndiffs, nsdiffs ,
                              feat_stl 
                              # , feature_set(pkgs = 'feasts') 
                             ) , 
                          s.window =  12 ) %>%
         mutate( sd / mean )
  ) %>%
  bind_rows()
  
  # View( f )
 
# stl features
#  this fails...
# train.4 %>% filter( orgUnit == ou[i] ) %>%
#   features( confCases, feat_stl, s.window = 'periodic' )
 
```

## fit 

```{r train4.fit, eval= FALSE }

# Run and save output interactively

## How to decipher which have a model???
i = 1:length( ou )

# special cases...
# i = which( ou =='AlGl0pjBwFS' )

if ( length( i ) < 6 ){
   train.4  %>%
    filter( orgUnit %in% ou[i] ) %>% autoplot()
}

fit.train.4 = train.4 %>%
  filter( orgUnit %in% ou[i] ) %>%
    model( 
          
          stl = STL( log(confCases + 1) ~ 
                       season( window = 13 ) + 
                       trend( window = 13 ) ) ,

          ets = ETS( log(confCases + 1)  ) ,
          

          ets_damped = ETS( log(confCases + 1)  ~ trend("Md") ,
                            restrict = FALSE ) ,
          
          seats = model(seats = feasts:::SEATS(log(confCases + 1))) ,
          
          arima = ARIMA( log(confCases + 1) ~ pdq(0:3,0:3,0:3) + PDQ(0:3,0:3,0:3)) ,
          
          fourier1 = ARIMA( log(confCases + 1) ~ 
                              fourier( K = 1 ) + PDQ(0,0,0)) ,
          
          fourier2 = ARIMA( log(confCases + 1) ~ 
                              fourier( K = 2 ) + PDQ(0,0,0))
           ) 


save( train.4, fit.train.4 , file = 'train4.Rda' )
# load( 'train4.Rda'  )
```

## decompose metrics 

```{r decomp_functions}

load( 'train4.Rda'  ) # train.4 and fit.train.4

  # get remainder formn stl 
  decomp_method = function( .mable ){
      
      modelCols = setdiff( colnames( .mable ) , key( .mable ) )
   
      # m = map( 1:length(models) , 
      #             ~ maple %>%
      #             pull( !! models[ .x ] ) %>% 
      #             map( 'fit' )  %>% 
      #             map( 'method' ) %>% unlist() 
      # ) 
      
      # model = .mable[, modelCols[.x] ]
      
      map_chr( 1:length( modelCols ) , 
               
               ~.mable[, modelCols[.x] ] %>% 
                 map( 1 ) %>%
                   map( 'fit' )  %>% 
                   map( class ) %>% unlist
              
           )
  
  }
  
  has_remainder = function( .mable ){

      ifelse( 
         any( 
         grepl('stl', decomp_method( .mable ) , 
               ignore.case = TRUE ) 
          ) 
      ,
              TRUE ,
              FALSE
      )
  }
  
  
  decomp_remainder = function( .mable ){
    if ( has_remainder( .mable ) ){
      .mable %>%
      pull( stl ) %>% 
      map( 'fit' )  %>% 
      map( 'decomposition' ) %>% 
        map( 'remainder' ) %>% 
        unlist()
    } else { NA }
  }
  
```

```{r model.4.metrics_examples}

i=1

# decomp method
dm = map( i , ~fit.train.4  %>% 
               filter( orgUnit %in% ou[ .x ] ) %>% 
                 summarise( 
                     method = decomp_method(.) %>%
                         paste( . , collapse = ", "), 
                     orgUnit = orgUnit
            )
        ) %>% 
  bind_rows()

dm 

# STL only
rsd = map( i , ~fit.train.4   %>% 
             select( orgUnit, stl ) %>% 
               filter( orgUnit %in% ou[ .x ] ) %>% 
                 summarise( 
                     sd_remainder = var( decomp_remainder(.) ,
                                         na.rm = TRUE )^.5 , 
                     orgUnit = orgUnit
            )
        ) %>% bind_rows()

```

```{r model.4.metrics}

summary( rsd$sd_remainder )

# non-STL !! LONG - convert to function with progress bar
rsd.all = map( i , ~fit.train.4  %>% 
               filter( orgUnit %in% ou[ .x ] ) %>% 
                select( -stl ) %>% 
                 augment() %>% 
                 as_tibble %>% 
                 group_by( orgUnit, .model ) %>%
                 summarise( 
                     sd_remainder = var( .resid ,
                                         na.rm = TRUE )^.5 
                     )
        ) %>% 
  bind_rows()

summary( rsd.all$sd_remainder )
rsd.all %>% group_by( .model ) %>% summarise( mean( sd_remainder , na.rm = TRUE ) )

# ou with highest remainder...
orgUnit.max.resid = rsd %>% arrange( desc( sd_remainder ) ) %>% 
    head(1) %>% pull( orgUnit ) 

  fit.train.4 %>% filter( orgUnit %in% orgUnit.max.resid ) %>%
    select( -stl ) %>%
    augment() %>% 
    autoplot(.fitted) +
      autolayer( train.4 %>% 
                   filter( orgUnit %in% orgUnit.max.resid ) %>%
                   mutate( .model = "Actual" ),
                 confCases ) + 
      labs( title =  'Highest Residual' , 
            subtitle = orgUnit.max.resid ) 

# ou with lowest remainder...
  orgUnit.min.resid = rsd %>% arrange(sd_remainder  ) %>% 
    head(1) %>% pull( orgUnit ) 

  fit.train.4 %>% filter( orgUnit %in% orgUnit.min.resid ) %>%
    select( -stl ) %>%
    augment() %>% 
    autoplot(.fitted) +
      autolayer( train.4 %>% 
                   filter( orgUnit %in% orgUnit.min.resid ) %>%
                   mutate( .model = "Actual" ),
                 confCases ) + 
      labs( title =  'Lowest Residual' , 
            subtitle = orgUnit.min.resid ) 


```

```{r, train4.stl_components, eval = FALSE}

# residual SD (from ETS -- try getting from other models...)
fit.train.4  %>% 
  filter( orgUnit %in% ou[i] ) %>%
  select( orgUnit , stl ) %>%
  # arrange( orgUnit ) %>%
  components(  ) %>% autoplot()
  as.tibble() %>% 
  group_by( orgUnit , .model ) %>% 
  dplyr::summarize( 
    n = n() ,
    mean = mean( remainder , na.rm = TRUE )  ,
    sd = var( remainder , na.rm = TRUE )^.5
    )

fit.train.4  %>% filter( orgUnit == ou[i] ) %>% 
  select( orgUnit , ets ) %>% augment()

  components(  ) %>% autoplot() 
```

```{r model_failed, eval=FALSE }

fail = fit.train.4  %>% 
  augment() %>% as.tibble() %>%
  group_by( .model , orgUnit ) %>%
  summarise( mean = mean( .fitted ) ) 

fail %>%
  summarise( n = n() , 
             fail = sum( is.na( mean ) ) 
             )

train.4 %>% as.tibble() %>%
  group_by( orgUnit ) %>%
  summarise( n = n() ,
             value = sum( !is.na( confCases ) )
             )

datasetTS %>% filter( orgUnit %in% 'a6SFNf7F567' )

```

## accuracy 

```{r model.4.accuracy_in_sample}
 
# fit.train.4 %>% unnest(fourier1 )

# fit.train.4 %>% count( fourier1$fit )
modelCols = setdiff( colnames( fit.train.4 ) , key( fit.train.4 ) ) %>% 
  setdiff( . , 'stl')

x = fit.train.4 %>% semi_join( long_enough[,]  , by = c("orgUnit") ) %>%
  select( -stl )

mase = point_accuracy_measures[6] 

accuracy.4.in_sample = accuracy(  x , measures = mase )

accuracy.4.in_sample %>% group_by(.model) %>%
  summarise( 
    series = n() ,
    trained = sum(!is.na( MASE ) ) ,
             mean = mean( MASE, na.rm = TRUE  )
             )

summary( accuracy.4.in_sample$MASE )

# accuracy( fit.train.4 %>% select( orgUnit, !! modelCols ) )

```

```{r forecast4.accuracy_out_of_sample, eval= FALSE }

# fore.train.4 = fit.train.4 %>%  
#   forecast( h = '13 months' )

# library(future)
# plan(multiprocess)



# bench::mark( 
#   
#   fore.train.4 = x[1:100,] %>% 
#   select( orgUnit, fourier1 ) %>% 
#   forecast( h = '13 months' )
#   
#   )

# List of forecasts
fore.train.4 =  map( 1:nrow( x ), ~x[ .x ,] %>% 
  select( orgUnit, fourier1 ) %>% 
  forecast( h = '13 months' ) ) 

# Print forecast with actual
.x = 1
fore.train.4[[.x]] %>% autoplot( level = c(80) ) +
  autolayer( datasetTS %>% filter( orgUnit %in% fore.train.4[[.x]]$orgUnit ) ,
             confCases ) +
  labs( title = .x )

# example multiple plots
fore.train.4[[1]] %>% autoplot( color = 'red' , level = c(80) ) + 
  autolayer( datasetTS %>% filter( orgUnit %in% fore.train.4[[1]]$orgUnit  ) ,
             confCases , color = 'red' , alpha = .5 ) +
  autolayer( fore.train.4[[2]] , alpha = .5 , level = c(80) ) +
  autolayer( datasetTS %>% filter( orgUnit %in% fore.train.4[[2]]$orgUnit ) ,
             confCases  , color = 'blue') 

# Accuray
## TODO: progress bar
fore.train.4.accuracy = map_df( 1:nrow( x ) , ~fore.train.4[[.x]] %>% 
  accuracy(
    datasetTS %>% filter( orgUnit %in% fore.train.4[[.x]]$orgUnit )
  )
)

summary(fore.train.4.accuracy$MASE)

save( fore.train.4 , fore.train.4.accuracy , file = 'fore.train.4.Rda' )
```

## Difference between actual and counterfactual 

```{r foretrain.4.difference}
load( 'fore.train.4.Rda' ) # fore.train.4 , fore.train.4.accuracy  

mape <- function(actual,pred){
           mape <- mean(abs((actual - pred)/actual), 
                        na.rm = TRUE )*100
           return (mape)
}

mpe <- function(actual,pred){
            pctActual = ifelse( pred == 0 , NA ,
                                (actual - pred)/pred
                                )
           mpe <- mean( pctActual  , 
                        na.rm = TRUE )*100
           return (mpe)
}

fore.actual.difference =  function( .x ){
  left_join( fore.train.4[[.x]] %>% 
               rename( predicted = confCases ) ,
             
             datasetTS %>% 
               filter( orgUnit %in% fore.train.4[.x][[1]]$orgUnit ) %>%
               rename( actual = confCases ) %>%
               select( orgUnit, month, actual ) , 
             
             by = c("orgUnit", "month")
             ) %>%
    as_tibble() %>%
    mutate( difference = actual - predicted ) %>%
    group_by( orgUnit ) %>%
    summarise( MPE = mpe( actual , predicted ) )
}

fore.train.4.difference = map_df( 1:nrow(x) , fore.actual.difference )

summary( fore.train.4.difference$MPE )

fore.train.4.difference %>% arrange( MPE ) %>% head
```

```{r fore.train.col.chart}

fore.train.4.difference %>% 
  arrange( MPE  ) %>% 
  mutate(orgUnit=factor(orgUnit, levels=orgUnit)) %>%
  ggplot() +
  geom_col(aes(y= MPE, x = orgUnit )) +
  scale_y_continuous(limits = c(-300,300)) +
  coord_flip() +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

```{r fore.train.extras, eval=FALSE}
biggestIncrease = fore.train.4.difference %>% arrange( desc( MPE ) ) %>% head(10) 

biggestDecline = fore.train.4.difference %>% arrange( MPE ) %>% head(10) 

# Examine extreme cases 
i = which( x$orgUnit %in% biggestIncrease$orgUnit[1])

fore.train.4[[i]] %>% autoplot( level = c(80) ) +
  autolayer( datasetTS %>% filter( orgUnit %in% fore.train.4[[i]]$orgUnit ) ,
             confCases ) +
  labs( title =  x$orgUnit[i] )

i = which( x$orgUnit %in% biggestDecline$orgUnit[6])

fore.train.4[[i]] %>% autoplot( level = c(80) ) +
  autolayer( datasetTS %>% filter( orgUnit %in% fore.train.4[[i]]$orgUnit ) ,
             confCases ) +
  labs( title =  x$orgUnit[i] )

datasetTS %>% filter( orgUnit %in% biggestDecline[1:3] ) %>% 
  select( orgUnit, month, confCases ) %>% autoplot(confCases)


```


#3 Impute data . ####

## check for outliers

Indication of outliers for orgUnits. Review all org units that reported >5,000 cases in any month.


```{r features_max}

sum_na = function(x ) sum(!is.na(x))

dts.features = datasetTS %>% 
  features(confCases, 
           c( n_obs = sum_na , feat_five_num ) 
           ) %>%
  
  mutate( high = ifelse( med > 0, max / med , NA) ) 

summary( dts.features$high )

dts.high = dts.features %>% filter( max>5000 ) 

datasetTS %>% 
  semi_join( dts.high , 
             by = c("orgUnit", "orgUnitName") ) %>%
autoplot( confCases )
```

```{r tsoutliers }

ous = datasetTS %>% 
  pull( orgUnit ) %>% unique()


confCases.ts = datasetTS %>% 
  filter( orgUnit %in% dts.high$orgUnit ) %>% 
  as_tibble() %>%
  select( orgUnit, month , confCases)  

# library( forecast )
# out = tsoutliers( confCases.ts , lambda = 'auto' )


# out = map( dts.high$orgUnit , 
#               ~tsoutliers::tso( confCases.ts %>% 
#                                   filter( orgUnit %in% .x  ) %>%
#                                   mutate( time = anytime::anytime(month) ) %>%
#                                   select( - month ) %>%
#                                   ts_ts
#                                   , 
#                                 tsmethod = "arima")
# )

# confCases.ts[ out$outliers$ind ]

### error in tsoutliers... ####

# outs = map( 11 , # length(dts.high$orgUnit) ,
#   ~ datasetTS %>% 
#   filter( orgUnit %in% dts.high$orgUnit[ .x ] ) %>% 
#   select( month , confCases)  %>%
#   as.ts %>%
#     tsoutliers::tso( log(.) , tsmethod = "auto.arima")
# )


long_enough = dts.features %>% filter( n_obs> 30 ) 

find_anomalies = function(.x , pb = NULL ){ 
  pb$tick()$print()
  
  datasetTS %>% 
    filter( !is.na( confCases) ) %>%
    semi_join( long_enough[.x, ] ,
                by = c("orgUnit", "orgUnitName") ) %>%
  select(  orgUnit , month , confCases)  %>%
  mutate( month = as.yearmon( month ) ) %>%
  as_tbl_time( index = month)  %>%
  time_decompose( confCases, method = "stl", frequency = 12 , trend = "auto" ,
                  message = FALSE ) %>%
  anomalize(remainder, method = "iqr", alpha = 0.01, max_anoms = 0.1 
            # , verbose=TRUE
            ) %>%
    mutate( orgUnit = long_enough$orgUnit[ .x ]  )
}

# TODO: add progress bar
pb = progress_estimated( nrow( long_enough ) )

out = map( 1:nrow( long_enough ) , 
           ~{ # print(.x)
             find_anomalies(.x , pb ) 
           }
) 

# out[[2]] %>% plot_anomaly_decomposition()

anomolies = map_df( out, ~.x %>% mutate( month = yearmonth( month) ) %>% 
                      as_tibble %>% 
                      filter( anomaly == 'Yes' )  %>%
                      select( orgUnit, everything() ) ) 

anomolies

nrow( anomolies ); count(anomolies, orgUnit )
```


```{r, brolgar }

d %>% as_tsibble( key = orgUnit , index = month) %>%
  sample_n_keys()

g = ggplot( d %>% as_tsibble( key = orgUnit , index = month)  ,
       aes(x = month ,
           y = confCases ,
           group = orgUnit)) +
  geom_line() +
  facet_sample() +
  labs(title='Selection of confCases reported by facility' , 
       subtitle = '12 random selections of 3 facilities')

g %>% plotly::ggplotly()
```



```{r dataset4_interpolate}

ous = unique( datasetTS$orgUnit )

datasetTS.models = datasetTS %>% 
  filter( orgUnit %in% ous ) %>%
  fill_gaps() %>%
  group_by( orgUnit ) %>% 
    model( 
    fourier1 = ARIMA( log(confCases + 1) ~ 
                              fourier( K = 1 ) + PDQ(0,0,0))
    ) 

model = map_chr( 1:length(ous) , 
                 ~datasetTS.models[.x,]$fourier1[[1]]$fit %>% class )

table( model )
has_model = model %in% 'ARIMA' 

itp = datasetTS.models %>%
  filter( has_model ) %>%
  interpolate( datasetTS ) %>%
  mutate( Interpolated = as.integer( round( confCases ) ) )


datasetTS.itp = itp %>% 
  select( -confCases ) %>%  
  left_join( datasetTS %>% 
               select( month, orgUnit, confCases ) %>%
               rename( Reported = confCases ) , 
             by = c("orgUnit", "month")
             )


#  see large outlier ... 
outOU = 'Dl6P4C549hR'

datasetTS.itp  %>% 
     filter( orgUnit %in% outOU ) %>% 
     gather( 'var' , 'val' , Interpolated, Reported ) %>%
       autoplot( val  ) +
       # labs( title = .x ) +
       # guides( color = FALSE ) +
       theme_bw( base_size = 10 )
     


```

```{r reNationalize }

confCases.0.itp = datasetTS.itp  %>% 
   summarise_if( is.numeric , sum , na.rm = TRUE ) %>%
   mutate( orgUnit = 'O' , orgUnitName = 'National' ) 
 
datasetTS.0.n =  datasetTS.itp %>% 
   summarise_if( is.numeric , ~sum(!is.na(.x))  ) %>%
   mutate( orgUnit = 'O' , orgUnitName = 'National' ) 

  
confCases.0.itp %>% 
       as_tibble() %>% 
       pivot_longer( c(Reported, Interpolated ), 
                     names_to = 'var' , values_to = 'val' ) %>%
       as_tsibble( index = month, 
                   key = c( var, orgUnit, orgUnitName ) ) %>%
       autoplot( val   ) +
       labs( title = 'confCases' ) +
       theme_bw( base_size = 10 )
    
  
```

```{r refit, eval=FALSE}

# Refit with national model
fit0.arima = fit0 %>% select( arima )

datasetTS.itp = map( ous[1:5], ~datasetTS %>% 
                       filter( orgUnit %in% .x ) %>%
                       refit( fit0.arima , . )
)
```

Model fit characteristics including characteristics of model, seasonal fit.  
 ETS, ARIMA., Fourier, Explanatory with ARIMA.  
 
 We evaluated impacted by 
 
 1. simulating a before-after cross-sectional study, choosing one month preceding and one month following bednet distribution.
 
 2. Using time-series models to forecast counterfactual data and compare with actual data.  Facilities were ranked by percent difference. The effect of bednets was estimated by conducting rank-sum test comparing facilites with new nets to those without.  